\chapter*{Aknowledgment}

Words cannot express my gratitude to Prof. Riadh Robbana, my academic supervisor, for his excellent mentorship not only throughout my thesis journey, but also my academic journey.

I am sincerely thankful to Prof. Manuel Bodirsky, my supervisor at TU Dresden, for
providing me with the invaluable opportunity to research . His expertise, insight and kindness boosted my learning
experience and expanded my horizons.

I am greatly thankful to Mr. Florian Starke, my co-supervisor at TU Dresden, for not only providing guidance, but also for being a great friend.

I would also like to express my recognition to my esteemed professors whose excellent courses and dedicated mentorship have equipped me with the knowledge and
skills necessary for this achievement. Their commitment to education has been instrumental in my academic growth.

I am indebted to the INSAT community for fostering an environment of
inspiration and growth. The collective spirit of students, faculty, and staff has continuously motivated me to strive for excellence.

Through this acknowledgment, I express my heartfelt gratitude to my friends who have been my pillars of strength.

Lastly, I would be remiss in not mentioning my family, my father Anouar, my mother Narjes and my two sisters Souha and Safa, Their belief in me has kept my spirits and motivation high during this process. I would also like to thank my cat Lara for all the entertainment and emotional support during the writing of this thesis.
\\
\\
\\
With heartfelt thanks,

Rami.




