\chapter*{Introduction}

\addcontentsline{toc}{chapter}{Introduction}

With the rising power of \acrfull{ai}, and especially \acrfull{dl}, \acrshort{ai} is being applied with a huge success to problems that were computationnally intractable.

In this report, we will focus on a family of games known as \acrfullpl{mpg}. It has been studied and analysed extensively with the hopes of finding an efficient general method to solve instances of that game. To the day of writing this report, no breakthroughs were being found. 

Instead, we will try to approximate solutions using a completely different method based on \acrfull{rl}, with an underlying \acrshort{dl} algorithm using \acrfull{sp} to refine its understanding of the game. Our methods will not only learn individual instances of the game, but also generalise to a whole class of \acrshortpl{mpg}

In the first chapter, we will give a presentation about the host insitute, an informal introduction to \acrshortpl{mpg} and talk about the state of the art in \acrshort{mpg} theory, its relation to other fields, and the existing \acrfull{ml} methods in related games.

In the second chapter, we will formalise \acrshort{mpg} using known results from the existing theory. Then, we will design and implement a \acrshort{mpg} library that is based on the theoretical frameworks. This library will be used by default in all the remaining of the project.

In the third chapter, we will generate two datasets of \acrshort{mpg} using efficient implementations of graph generation. We will then try to solve all instances of both datasets using our most efficient implementation of \acrshort{mpg} solver based on \acrfull{csp} methods.  We will use also a \acrfull{hpc} cluster to further minimise the execution time.

In the fourth chapter, we will analyse briefly both datasets with their annotations, and try to confirm empirically our hypotheses about \acrshort{mpg}. We will then design a \acrshort{gnn} model that conforms to some desired properties, based in our findings.

In the fifth and final chapter, we will implement a whole distributed \acrshort{rl} system that uses \acrfull{sp} that refines the learnable parameters of the model discussed in chapter \ref{chapter:ModelDesign}.  We will base the full architecture of the system in the famous Alpha Zero paper \cite{AlphaZero}. Finally, we will deploy the system in a large \acrshort{hpc} cluster and execute it for refining the model.

